---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.7.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---


# Quiz: Reading and Summarizing Data


## Imports


```{r}
suppressPackageStartupMessages(library(R.utils))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(readxl))
```


## First Dataset


```{r}
# Load the dataset:
part_matter <- suppressMessages(read_csv("daily_SPEC_2014.csv.bz2"))

# Print the first 20 rows:
head(part_matter, 20L)
```


## Question 1

What is the average "Arithmetic Mean" for "Bromine PM2.5 LC" in the state of Wisconsin?


```{r}
# Compute the desired average:
amean <- part_matter["Arithmetic Mean"]
fltr_pname <- part_matter["Parameter Name"] == "Bromine PM2.5 LC"
fltr_state <- part_matter["State Name"] == "Wisconsin"
avg_lvl <- mean(amean[fltr_pname & fltr_state], na.rm = TRUE)
printf("Average = %.6f\n", avg_lvl)
```


## Question 2

Which constituent "Parameter Name" has the highest average level?


```{r}
# Use the column "Parameter Name" to group the data:
by_pname <- group_by_at(part_matter, vars("Parameter Name"))

# Compute the average level for each group:
avg_lvl <- summarise_at(by_pname, vars("Arithmetic Mean"), mean, na.rm = TRUE)

pname <- pull(avg_lvl, 1)
avlvl <- pull(avg_lvl, 2)

# Constituent with the highest average level:
const <- pname[avlvl == max(avlvl[grepl("PM2.5", pname)])]
printf("Constituent with the highest average level = %s\n", const)
```


## Question 3

Which monitoring site has the highest average level of "Sulfate PM2.5 LC"
across all time?


```{r}
# Use the columns "State Code", "County Code" and "Site Num" to group the data:
by_msite <- group_by_at(part_matter, vars("State Code", "County Code", "Site Num"))

# Extract the data corresponding to "Sulfate PM2.5 LC":
by_msite_sulf <- filter_at(by_msite, vars("Parameter Name"), ~ .x == "Sulfate PM2.5 LC")

# Compute the average level for each group:
avg_lvl <- summarise_at(by_msite_sulf, vars("Arithmetic Mean"), mean, na.rm = TRUE)

# Monitoring site with the highest average level:
tmp <- filter_at(avg_lvl, vars("Arithmetic Mean"), ~ .x == max(avg_lvl["Arithmetic Mean"]))
printf("Monitoring site with the highest average level:\n")
printf("State Code  - %s\n", pull(tmp, 1))
printf("County Code - %s\n", pull(tmp, 2))
printf("Site Num    - %s\n", pull(tmp, 3))
```


## Question 4

What is the absolute difference in the average levels of "EC PM2.5 LC TOR"
between the states California and Arizona, across all time and all monitoring
sites?


```{r}
fltr_pname <- part_matter["Parameter Name"] == "EC PM2.5 LC TOR"

# Compute the average level for Arizona:
fltr_state <- part_matter["State Name"] == "Arizona"
avg_lvl_az <- mean(amean[fltr_pname & fltr_state], na.rm = TRUE)

# Compute the average level for California:
fltr_state <- part_matter["State Name"] == "California"
avg_lvl_ca <- mean(amean[fltr_pname & fltr_state], na.rm = TRUE)

# Compute the absolute difference:
abs_diff <- abs(avg_lvl_az - avg_lvl_ca)
printf("Absolute difference = %.6f\n", abs_diff)
```


## Question 5

What is the median level of "OC PM2.5 LC TOR" in the western United States,
across all time? Define western as any monitoring location that has a longitude
less than -100.


```{r}
# Compute the median level:
fltr_long <- part_matter["Longitude"] < -100
fltr_pname <- part_matter["Parameter Name"] == "OC PM2.5 LC TOR"
median_lvl <- median(amean[fltr_long & fltr_pname], na.rm = TRUE)
printf("Median level = %.4f\n", median_lvl)
```


## Second Dataset


```{r}
# Load the dataset:
col_types <- c(rep("text", times = 3), rep("guess", times = 25))
sites <- read_excel("aqs_sites.xlsx", col_types = col_types)

# Print the first 20 rows:
head(sites, 20L)
```


## Question 6

How many monitoring sites are labelled as both "RESIDENTIAL" for "Land Use" and
"SUBURBAN" for "Location Setting"?


```{r}
# Compute the number of monitoring sites:
fltr_loc <- sites["Location Setting"] == "SUBURBAN"
fltr_luse <- sites["Land Use"] == "RESIDENTIAL"
sites_count <- sum(fltr_loc & fltr_luse, na.rm = TRUE)
printf("Number of monitoring sites = %d\n", sites_count)
```


## Question 7

What is the median level of "EC PM2.5 LC TOR" amongst monitoring sites that are
labelled as both "RESIDENTIAL" and "SUBURBAN" in the eastern US, where eastern
is defined as longitude greater than or equal to -100?


```{r}
# Get the relevant data from the 1st dataset:
fltrd_1 <- filter_at(part_matter, vars("Parameter Name"), ~ .x == "EC PM2.5 LC TOR")
fltrd_1 <- fltrd_1[c("State Code", "County Code", "Site Num", "Arithmetic Mean")]
fltrd_1 <- mutate_at(fltrd_1, vars("State Code", "County Code", "Site Num"), as.numeric)

# Get the relevant data from the 2nd dataset:
fltrd_2 <- filter(
    sites,
    .data[["Land Use"]] == "RESIDENTIAL",
    .data[["Location Setting"]] == "SUBURBAN",
    .data[["Longitude"]] >= -100
)
fltrd_2 <- fltrd_2[c("State Code", "County Code", "Site Number")]
colnames(fltrd_2)[3] <- "Site Num"
fltrd_2 <- mutate_at(fltrd_2, vars("State Code", "County Code", "Site Num"), as.numeric)

# Compute the median level:
jds <- inner_join(fltrd_1, fltrd_2, by = c("State Code", "County Code", "Site Num"))
median_lvl <- median(jds[["Arithmetic Mean"]], na.rm = TRUE)
printf("Median level = %.4f\n", median_lvl)
```


## Question 8

Amongst monitoring sites that are labeled as "COMMERCIAL" for "Land Use", which
month of the year has the highest average levels of "Sulfate PM2.5 LC"?


```{r}
# Get the relevant data from the 1st dataset:
fltrd_1 <- filter_at(part_matter, vars("Parameter Name"), ~ .x == "Sulfate PM2.5 LC")
fltrd_1 <- fltrd_1[c("State Code", "County Code", "Site Num", "Arithmetic Mean", "Date Local")]
fltrd_1 <- mutate_at(fltrd_1, vars("State Code", "County Code", "Site Num"), as.numeric)
fltrd_1 <- mutate_at(fltrd_1, vars("Date Local"), ~ format(.x, "%m"))
colnames(fltrd_1)[5] <- "Month"

# Get the relevant data from the 2nd dataset:
fltrd_2 <- filter_at(sites, vars("Land Use"), ~ .x == "COMMERCIAL")
fltrd_2 <- fltrd_2[c("State Code", "County Code", "Site Number")]
colnames(fltrd_2)[3] <- "Site Num"
fltrd_2 <- suppressWarnings(mutate_at(fltrd_2, vars("State Code", "County Code", "Site Num"), as.numeric))

# Month with the highest average level:
jds <- inner_join(fltrd_1, fltrd_2, by = c("State Code", "County Code", "Site Num"))
avg_lvl <- summarise_at(group_by_at(jds, vars("Month")), vars("Arithmetic Mean"), mean, na.rm = TRUE)
mon <- pull(filter_at(avg_lvl, vars("Arithmetic Mean"), ~ .x == max(avg_lvl["Arithmetic Mean"])), 1)
printf("Month with the highest average level = %s\n", mon)
```


## Question 9

Take a look at the data for the monitoring site identified by State Code 6,
County Code 65, and Site Number 8001 (this monitor is in California). At this
monitor, for how many days is the sum of "Sulfate PM2.5 LC" and "Total Nitrate
PM2.5 LC" greater than 10?

For each of the chemical constituents, there will be some dates that have
multiple "Arithmetic Mean" values at this monitoring site. When there are
multiple values on a given date, take the average of the constituent values for
that date.


```{r}
# Get the relevant data from the 1st dataset:
fltrd_1 <- filter(
    part_matter,
    as.numeric(.data[["State Code"]]) == 6,
    as.numeric(.data[["County Code"]]) == 65,
    as.numeric(.data[["Site Num"]]) == 8001,
    .data[["Parameter Name"]] == "Sulfate PM2.5 LC" | .data[["Parameter Name"]] == "Total Nitrate PM2.5 LC"
)
fltrd_1 <- fltrd_1[c("Date Local", "Parameter Name", "Arithmetic Mean")]

# Sulfate:
sulf <- filter_at(fltrd_1, vars("Parameter Name"), ~ .x == "Sulfate PM2.5 LC")
sulf <- sulf[c("Date Local", "Arithmetic Mean")]
colnames(sulf)[2] <- "Sulfate"
sulf <- summarise_at(group_by_at(sulf, vars("Date Local")), vars("Sulfate"), mean, na.rm = TRUE)

# Total Nitrate:
tot_nit <- filter_at(fltrd_1, vars("Parameter Name"), ~ .x == "Total Nitrate PM2.5 LC")
tot_nit <- tot_nit[c("Date Local", "Arithmetic Mean")]
colnames(tot_nit)[2] <- "Total Nitrate"
tot_nit <- summarise_at(group_by_at(tot_nit, vars("Date Local")), vars("Total Nitrate"), mean, na.rm = TRUE)

# Compute the number of days:
jds <- inner_join(sulf, tot_nit, by = "Date Local")
days <- nrow(filter(jds, .data[["Sulfate"]] + .data[["Total Nitrate"]] > 10))
printf("Number of days = %d\n", days)
```


## Question 10

Which monitoring site in the dataset has the highest correlation between
"Sulfate PM2.5 LC" and "Total Nitrate PM2.5 LC" across all dates? Identify the
monitoring site by its State, County, and Site Number code.


```{r}
# Get the relevant data from the 1st dataset:
fltrd_1 <- filter_at(
    part_matter, vars("Parameter Name"),
    ~ .x == "Sulfate PM2.5 LC" | .x == "Total Nitrate PM2.5 LC"
)
fltrd_1 <- fltrd_1[c("State Code", "County Code", "Site Num", "Date Local", "Parameter Name", "Arithmetic Mean")]

# Sulfate:
sulf <- filter_at(fltrd_1, vars("Parameter Name"), ~ .x == "Sulfate PM2.5 LC")
sulf <- sulf[c("State Code", "County Code", "Site Num", "Date Local", "Arithmetic Mean")]
colnames(sulf)[5] <- "Sulfate"
sulf <- summarise_at(
    group_by_at(
        sulf, vars("State Code", "County Code", "Site Num", "Date Local")
    ),
    vars("Sulfate"), mean, na.rm = TRUE
)

# Total Nitrate:
tot_nit <- filter_at(fltrd_1, vars("Parameter Name"), ~ .x == "Total Nitrate PM2.5 LC")
tot_nit <- tot_nit[c("State Code", "County Code", "Site Num", "Date Local", "Arithmetic Mean")]
colnames(tot_nit)[5] <- "Total Nitrate"
tot_nit <- summarise_at(
    group_by_at(
        tot_nit, vars("State Code", "County Code", "Site Num", "Date Local")
    ),
    vars("Total Nitrate"), mean, na.rm = TRUE
)

# Monitoring site with the highest correlation:
jds <- inner_join(sulf, tot_nit, by = c("State Code", "County Code", "Site Num", "Date Local"))
corrs <- summarise(
    jds, Correlations = cor(.data[["Sulfate"]], .data[["Total Nitrate"]]), .groups = "drop"
)
tmp <- filter_at(corrs, vars("Correlations"), ~ .x == max(corrs["Correlations"]))
printf("Monitoring site with the highest correlation:\n")
printf("State Code  - %s\n", pull(tmp, 1))
printf("County Code - %s\n", pull(tmp, 2))
printf("Site Num    - %s\n", pull(tmp, 3))
```
