---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.7.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---


# Quiz: Reading and Summarizing Data


## Imports


```{r}
suppressPackageStartupMessages(library(R.utils))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(readxl))
```


## First Dataset


```{r}
# Load the dataset:
part_matter <- suppressMessages(read_csv("daily_SPEC_2014.csv.bz2"))
```


```{r}
# Print the first 20 rows:
head(part_matter, 20L)
```


```{r}
# Print the column names:
colnames(part_matter)
```


## Question 1

What is the average "Arithmetic Mean" for "Bromine PM2.5 LC" in the state of Wisconsin?


```{r}
# Helper function:
compute_avg_lvl <- function(constituent, state) {
    amean <- part_matter["Arithmetic Mean"]
    fltr_pname <- part_matter["Parameter Name"] == constituent
    fltr_state <- part_matter["State Name"] == state
    avg_lvl <- mean(amean[fltr_pname & fltr_state], na.rm = TRUE)
    return(avg_lvl)
}
```


```{r}
# Compute and print the desired average:
avg_lvl <- compute_avg_lvl("Bromine PM2.5 LC", "Wisconsin")
printf("Average = %.6f\n", avg_lvl)
```


## Question 2

Which constituent "Parameter Name" has the highest average level?


```{r}
# Use the column "Parameter Name" to group the data:
by_pname <- group_by_at(part_matter, vars("Parameter Name"))

# Compute the average level for each group:
avg_lvl <- summarise_at(by_pname, vars("Arithmetic Mean"), mean, na.rm = TRUE)

# Highest average level:
pname <- pull(avg_lvl, 1)
amean <- pull(avg_lvl, 2)
fltr <- grepl("PM2.5", pname)
high_avg_lvl <- max(amean[fltr])

# Corresponding constituent:
fltr <- amean == high_avg_lvl
const <- pname[fltr]
printf("Constituent with the highest average level = %s\n", const)
```


## Question 3

Which monitoring site has the highest average level of "Sulfate PM2.5 LC"
across all time?


```{r}
# Use the columns "State Code", "County Code" and "Site Num" to group the data:
by_msite <- group_by_at(
    part_matter, vars("State Code", "County Code", "Site Num")
)

# Extract the data corresponding to "Sulfate PM2.5 LC":
by_msite_sulf <- filter_at(
    by_msite, vars("Parameter Name"), ~ .x == "Sulfate PM2.5 LC"
)

# Compute the average level for each group:
avg_lvl <- summarise_at(
    by_msite_sulf, vars("Arithmetic Mean"), mean, na.rm = TRUE
)

printf("Monitoring site with the highest average level:\n")
tmp <- filter_at(
    avg_lvl, vars("Arithmetic Mean"), ~ .x == max(avg_lvl["Arithmetic Mean"])
)
printf("State Code  - %s\n", pull(tmp, 1))
printf("County Code - %s\n", pull(tmp, 2))
printf("Site Num    - %s\n", pull(tmp, 3))
```


## Question 4

What is the absolute difference in the average levels of "EC PM2.5 LC TOR"
between the states California and Arizona, across all time and all monitoring
sites?


```{r}
# Compute the average level for Arizona:
avg_lvl_az <- compute_avg_lvl("EC PM2.5 LC TOR", "Arizona")

# Compute the average level for California:
avg_lvl_ca <- compute_avg_lvl("EC PM2.5 LC TOR", "California")

# Compute the absolute difference:
abs_diff <- abs(avg_lvl_az - avg_lvl_ca)
printf("Absolute difference = %.6f\n", abs_diff)
```


## Question 5

What is the median level of "OC PM2.5 LC TOR" in the western United States,
across all time? Define western as any monitoring location that has a Longitude
LESS THAN -100.


```{r}
# Compute the median level:
amean <- part_matter["Arithmetic Mean"]
fltr_long <- part_matter["Longitude"] < -100
fltr_pname <- part_matter["Parameter Name"] == "OC PM2.5 LC TOR"
median_lvl <- median(amean[fltr_long & fltr_pname], na.rm = TRUE)
printf("Median level = %.4f\n", median_lvl)
```


## Second Dataset


```{r}
# Load the dataset:
sites <- suppressWarnings(read_excel("aqs_sites.xlsx"))
```


```{r}
# Print the first 20 rows:
head(sites, 20L)
```


```{r}
# Print the column names:
colnames(sites)
```


## Question 6

How many monitoring sites are labelled as both RESIDENTIAL for "Land Use" and
SUBURBAN for "Location Setting"?


```{r}
# Compute the number of monitoring sites:
fltr_loc <- sites["Location Setting"] == "SUBURBAN"
fltr_luse <- sites["Land Use"] == "RESIDENTIAL"
sites_count <- sum(fltr_loc & fltr_luse, na.rm = TRUE)
printf("Number of monitoring sites = %d\n", sites_count)
```


## Question 7

What is the median level of "EC PM2.5 LC TOR" amongst monitoring sites that are
labelled as both "RESIDENTIAL" and "SUBURBAN" in the eastern US, where eastern
is defined as longitude greater than or equal to -100?


```{r}
# x:
```
